%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape
Non solo nella musica, ma anche nella vita \\
il vero spettacolo è ascoltare. \\ \medskip
Not only in music, but also in life \\
true wonder lies in listening. \\ \bigskip
-- Stefano Bollani
}
\end{flushright}



\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}
First of all, I would like to thank my supervisor Perfecto Herrera for encouraging me to pursue such an unusual kind of research. I always felt that not enough people are willing to put themselves in the middle ground between the engineering and the humanistic approaches to knowledge, so I am happy to having been given the opportunity to fill in this gap. Also, I know in many occasions I haven't been the most diligent student, so thanks for bearing with me until the end.

A great amount of support has been provided by Xavier Favory at the \acsfont{MTG}. Thanks to him for the inspiring insights about many technicalities involved, which definitely helped me orient myself in the world of \acsfont{NLP}. I would have not even started my work if it wasn't for his help.

I am hugely grateful to the awesome people in the Sound and Music Computing class of 2018, for the continuous mutual support, the tough moments spent together, the inspiring discussions, and all the beers.

Finally, everything would have been way harder if it wasn't for my family -- no, my two families: the one I left in Italy, and the one that welcomed me in Barcelona. Muchissimo amor to Dani, Joe, Minz and Tessy, I still hope one day you will understand how wrong pineapple on pizza is.

\bigskip

\endgroup



